---
title: "Inference with Categorical Data"
author: Dr. Gilbert
format: revealjs
date: today
date-format: long
theme: serif
incremental: true
fontsize: 20pt
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(patchwork)
library(countdown)

options(kable_styling_bootstrap_options = c("hover", "striped"))

theme_set(theme_bw(base_size = 32))

pop_prop <- 0.86
num_obs <- 17
new_obs <- 113
num_samps <- 100
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}
```

## The Highlights

+ Reminder on Proportions
+ Sampling from a Population
+ Sampling Variation
+ Repeated Sampling and Proportion Estimates
    + Sampling Error
+ Another item...
+ Other additional items...
+ Example Problems (confidence intervals and hypothesis tests)
+ Summary

## Reminder on Proportions

. . . 

Proportions are a statistic for summarizing categorical data

. . . 

Proportions measure part of a whole, and are typically represented as a decimal value between 0 and 1 -- alternatives include fractions or percentages.

. . .

Proportions are used to measure the relative size of a group or category resulting from *grouping questions* like:

+ Are you a smoker?
+ Do you identify politically as *conservative*, *moderate*, *liberal*, or *unsure*?
+ Is the type of diet *carnivore*, *herbivore*, *omnivore*, *insectivore*, or *frugivore*?

## Sampling Variation

. . . 

When we collect sample data, we get an estimate for the population parameter.

. . . 

We do this because we want to know the population parameter, even though we are unable to measure it directly.

. . . 

I'm going to take a random sample of `r num_obs` individuals from a voting population considering the fictitious *Proposition B-52*. The results of the individuals' voting plans appear below.

. . .

```{r}
my_sample_1 <- sample(c("In-Favor", "Against"), 
                    size = num_obs, 
                    prob = c(pop_prop, 1 - pop_prop),
                    replace = TRUE)

my_prop_est_1 <- mean(my_sample_1 == "In-Favor")

my_sample_1
```

. . . 

The estimated proportion in favor of *Proposition B-52* is about `r round(my_prop_est_1, 3)`

. . . 

Is the population proportion in favor of *Proposition B-52* equal to `r round(my_prop_est_1, 3)`?

. . . 

> No, but the sample provides an *estimate* for the population proportion.

## Sampling Variation

Here's another random sample of `r num_obs` individuals' voting preference on *Proposition B-52*.

. . . 

```{r}
my_sample_2 <- sample(c("In-Favor", "Against"), 
                    size = num_obs, 
                    prob = c(pop_prop, 1 - pop_prop),
                    replace = TRUE)

my_prop_est_2 <- mean(my_sample_2 == "In-Favor")

my_sample_2
```

. . . 

The estimated proportion from this sample is `r round(my_prop_est_2, 3)`

. . . 

Okay, so what can the results from a random sample actually tell us?

## Sampling Variation and Sampling Error

. . . 

::::{.columns}

:::{.column width="40%"}

The result from our first sample appears in the plot on the right.

:::

:::{.column width="60%"}

```{r}
#| fig.height: 7

my_samples <- tibble(
  sample_id = 1:num_samps,
  prop_est = c(my_prop_est_1, rep(NA, num_samps - 1))
)

my_samples %>%
  ggplot() + 
  geom_point(aes(x = prop_est,
                 y = sample_id)) + 
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(0, 101)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

:::

::::

## Sampling Variation and Sampling Error

::::{.columns}

:::{.column width="40%"}

We've added in the estimate from our second sample as well.

:::

:::{.column width="60%"}

```{r}
#| fig.height: 7

my_samples <- tibble(
  sample_id = 1:num_samps,
  prop_est = c(my_prop_est_1, my_prop_est_2, rep(NA, num_samps - 2))
)

my_samples %>%
  ggplot() + 
  geom_point(aes(x = prop_est,
                 y = sample_id)) + 
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(0, 101)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

:::

::::

## Sampling Variation and Sampling Error

::::{.columns}

:::{.column width="40%"}

Here's the result of a third random sample of `r num_obs` individuals.

:::

:::{.column width="60%"}

```{r}
#| fig.height: 7

my_sample_3 <- sample(c("In-Favor", "Against"), 
                    size = num_obs, 
                    prob = c(pop_prop, 1 - pop_prop),
                    replace = TRUE)

my_prop_est_3 <- mean(my_sample_3 == "In-Favor")

my_samples <- tibble(
  sample_id = 1:num_samps,
  prop_est = c(my_prop_est_1, my_prop_est_2, my_prop_est_3, rep(NA, num_samps - 3))
)

my_samples %>%
  ggplot() + 
  geom_point(aes(x = prop_est,
                 y = sample_id)) + 
  coord_cartesian(xlim = c(0, 1),
                  ylim = c(0, 101)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

:::

::::

## Sampling Variation and Sampling Error

::::{.columns}

:::{.column width="40%"}

Let's take `r num_samps - 3` more random samples and plot the results from all of these random samples containing `r num_obs` voters each... 

:::

:::{.column width="60%"}

```{r}
#| fig.height: 7

toy_data <- tibble(
  sample = 1:num_samps,
  samp_prop = c(my_prop_est_1, my_prop_est_2, my_prop_est_3, rep(NA, num_samps - 3))
                   )

for(i in 4:num_samps){
  toy_data[i, "samp_prop"] <- rbinom(1, num_obs, pop_prop)/num_obs
}

toy_data %>%
  ggplot() + 
  # geom_vline(xintercept = pop_prop, 
  #            linetype = "dashed") + 
  geom_point(aes(x = samp_prop, 
                 y = sample)) +
  coord_cartesian(xlim = c(0, 1)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

:::

::::

. . . 

What information can we take away from these `r num_samps` random samples?

## Sampling Variation and Sampling Error

. . . 

A sample of 17 individuals is quite small and leaves lots of uncertainty.

. . . 

Let's take samples of `r new_obs` individuals instead

::::{.columns}

:::{.column width="40%"}

Here are `r num_samps` random samples of `r new_obs` voters each.

:::

:::{.column width="60%"}

```{r}
#| fig.height: 7

toy_data <- tibble(
  sample = 1:num_samps,
  samp_prop = rep(NA, num_samps),
  samp_prop_norm = rep(NA, num_samps)
                   )

for(i in 1:num_samps){
  toy_data[i, "samp_prop"] <- rbinom(1, new_obs, pop_prop)/new_obs
  toy_data[i, "samp_prop_norm"] <- rnorm(1, mean = pop_prop, 
                                         sd = sqrt(pop_prop*(1 - pop_prop)/new_obs))
}

toy_data %>%
  ggplot() + 
  # geom_vline(xintercept = pop_prop, 
  #            linetype = "dashed") + 
  geom_point(aes(x = samp_prop, 
                 y = sample)) +
  # geom_point(aes(x = samp_prop_norm,
  #                y = sample),
  #            color = "red",
  #            alpha = 0.2) + 
  coord_cartesian(xlim = c(0, 1)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

:::

::::

. . . 

What do we feel comfortable claiming now?

## Sampling Variation and Sampling Error (Summary, So Far...)

. . . 

Each sample provides us with a *summary statistic* (a sample proportion) which serves as an estimate for the population proportion.

. . . 

Each sample results in a different *summary statistic* -- this is referred to as *sampling variation*

## Sampling Variation and Sampling Error

. . . 

Let's revisit those `r num_samps` random samples of `r new_obs` individuals.

. . .

```{r}
toy_data %>%
  ggplot() + 
  # geom_vline(xintercept = pop_prop, 
  #            linetype = "dashed") + 
  geom_point(aes(x = samp_prop, 
                 y = sample)) +
  # geom_point(aes(x = samp_prop_norm,
  #                y = sample),
  #            color = "red",
  #            alpha = 0.2) + 
  coord_cartesian(xlim = c(0, 1)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

## Sampling Variation and Sampling Error

Let's revisit those `r num_samps` random samples of `r new_obs` individuals.

```{r}
toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, 
             linetype = "dashed") + 
  geom_point(aes(x = samp_prop, 
                 y = sample)) +
  # geom_point(aes(x = samp_prop_norm,
  #                y = sample),
  #            color = "red",
  #            alpha = 0.2) + 
  coord_cartesian(xlim = c(0, 1)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

. . . 

That black dashed line is at the underlying population proportion (`r pop_prop`)

## Sampling Variation and Sampling Error

Let's revisit those `r num_samps` random samples of `r new_obs` individuals.

```{r}
toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, 
             linetype = "dashed") + 
  geom_point(aes(x = samp_prop, 
                 y = sample)) +
  # geom_point(aes(x = samp_prop_norm,
  #                y = sample),
  #            color = "red",
  #            alpha = 0.2) + 
  geom_segment(aes(x = samp_prop, xend = pop_prop, y = sample, yend = sample),
               color = "red") +
  coord_cartesian(xlim = c(0, 1)) + 
  labs(
    x = "Sample Proportion",
    y = "Sample Number"
  )
```

. . . 

Now, these red horizontal lines represent the *sampling error* -- how far the sample statistic falls from the true population proportion.

## Distribution of Sampling Errors

. . . 

Let's switch our view here and look at the distribution of sampling errors.

. . . 

```{r}
toy_data %>%
  mutate(samp_errors = samp_prop - pop_prop) %>%
  ggplot() + 
  geom_histogram(aes(x = samp_errors, y = after_stat(density)),
                 color = "black",
                 fill = "purple",
                 bins = 15) + 
  geom_density(aes(x = samp_errors),
               fill = "purple",
               alpha = 0.4) + 
  labs(
    x = "Sampling Error",
    y = "Density"
  )
```

. . .

What do you notice?

## Distribution of Sampling Errors

. . . 

In case you aren't convinced, here's the distribution of sampling errors from 10,000 random samples.

. . . 

```{r}
many_samples <- tibble(
  sample_prop = rbinom(1e3, new_obs, pop_prop)/new_obs
) %>%
  mutate(samp_errors = sample_prop - pop_prop) 

many_samples %>%
  ggplot() + 
  geom_histogram(aes(x = samp_errors, y = after_stat(density)),
                 color = "black",
                 fill = "purple") + 
  geom_density(aes(x = samp_errors),
               fill = "purple",
               alpha = 0.4) + 
  labs(
    x = "Sampling Error",
    y = "Density"
  )
```

## Sampling Variation and Sampling Error

. . . 

Each sample provides us with a *summary statistic* (a sample proportion) which serves as an estimate for the population proportion.

. . . 

Each sample results in a different *summary statistic* -- this is referred to as *sampling variation*.

. . . 

Each sample's *summary statistic* has some deviation from the *population proportion* -- that deviation is called the *sampling error*.

. . . 

The distribution of sampling errors is approximately normal$^*$

## When Are Sampling Errors Approximately Normal?

. . . 

It's not actually always the case that the sampling errors will be nearly normally distributed.

. . . 

Let's consider three example scenarios -- one in which we took samples of size 10, the next with samples of size 40, and finally, with samples of size 100.

```{r}
#| fig.height: 5

toy_data <- tibble(
  sample_id = 1:100,
  small_samples = rbinom(100, 15, pop_prop)/15,
  medium_samples = rbinom(100, 40, pop_prop)/40,
  large_samples = rbinom(100, 100, pop_prop)/100
)

p1 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = small_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = small_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 15)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 11)

p2 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = medium_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = medium_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 40)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 11)

p3 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = large_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = large_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 100)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 11)

p1 + p2 + p3
```

## When Are Sampling Errors Approximately Normal?

It's not actually always the case that the sampling errors will be nearly normally distributed.

Let's consider three example scenarios -- one in which we took samples of size 10, the next with samples of size 40, and finally, with samples of size 100.

::::{.columns}

:::{.column width="65%"}

```{r}
#| fig.height: 5

toy_data <- tibble(
  sample_id = 1:100,
  small_samples = rbinom(100, 15, pop_prop)/15,
  medium_samples = rbinom(100, 40, pop_prop)/40,
  large_samples = rbinom(100, 100, pop_prop)/100
)

p1 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = small_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = small_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 15)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 13)

p2 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = medium_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = medium_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 40)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 13)

p3 <- toy_data %>%
  ggplot() + 
  geom_vline(xintercept = pop_prop, linetype = "dashed") + 
  geom_segment(aes(x = pop_prop, xend = large_samples, 
                   y = sample_id, yend = sample_id),
               color = "red") +
  geom_point(aes(x = large_samples, y = sample_id)) + 
  labs(
    title = " Sampling Errors (n = 100)",
    x = "Sample Proportion",
    y = "Sample Number") + 
  coord_cartesian(xlim = c(0, 1)) + 
  theme_bw(base_size = 13)

p1 + p2 + p3
```

:::

:::{.column width="35%"}

```{r}
#| fig.height: 10

toy_data <- tibble(
  sample_id = 1:1e3,
  small_samples = rbinom(1e3, 15, pop_prop)/15,
  medium_samples = rbinom(1e3, 40, pop_prop)/40,
  large_samples = rbinom(1e3, 100, pop_prop)/100
)

toy_data <- toy_data %>%
  mutate(
    samp_err_small = small_samples - pop_prop,
    samp_err_medium = medium_samples - pop_prop,
    samp_err_large = large_samples - pop_prop
  )

p1 <- toy_data %>%
  ggplot() + 
  geom_histogram(aes(x = samp_err_small,
                     y = after_stat(density)),
                 color = "black",
                 fill = "purple") +
  geom_density(aes(x = samp_err_small),
               fill = "purple",
               alpha = 0.75) +
  labs(
    title = " Sampling Errors (n = 15)",
    x = "Sampling Error",
    y = "") +  
  theme_bw(base_size = 22)

p2 <- toy_data %>%
  ggplot() + 
  geom_histogram(aes(x = samp_err_medium,
                     y = after_stat(density)),
                 color = "black",
                 fill = "purple") +
  geom_density(aes(x = samp_err_medium),
               fill = "purple",
               alpha = 0.75) +
  labs(
    title = " Sampling Errors (n = 40)",
    x = "Sampling Error",
    y = "") + 
  theme_bw(base_size = 22)

p3 <- toy_data %>%
  ggplot() + 
  geom_histogram(aes(x = samp_err_large,
                     y = after_stat(density)),
                 color = "black",
                 fill = "purple") +
  geom_density(aes(x = samp_err_large),
               fill = "purple",
               alpha = 0.75) +
  labs(
    title = " Sampling Errors (n = 100)",
    x = "Sampling Error",
    y = "") + 
  theme_bw(base_size = 22)

p1 / p2 / p3
```

:::

::::

. . . 

**Rule of Thumb:** The 

## Next Time...

<center><br/> Discrete Probability and Simulation Lab</center>



