---
title: "Practicing Inference with Categorical Data"
subtitle: "One- and Two-Proportion Inference"
author: Dr. Gilbert
format: revealjs
date: today
date-format: long
theme: serif
incremental: true
fontsize: 20pt
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(patchwork)
library(countdown)

options(kable_styling_bootstrap_options = c("hover", "striped"))

theme_set(theme_bw(base_size = 32))

pop_prop <- 0.74
num_obs <- 17
new_obs <- 113
num_samps <- 100

set.seed(8202024)
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}

a {
  color: purple;
}

a:link {
  color: purple;
}

a:visited {
  color: purple;
}
```

## Reminder: Framework for Constructing Confidence Intervals

**General Strategy:** Each time we seek to Construct a confidence interval, we will...

1. Read the scenario very carefully
2. Determine what it is we are trying to estimate -- a mean? a proportion? a difference in means or proportions?
3. Recall the general "formula" for the *confidence interval*

. . .

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot S_E}$$

4. Re-read the scenario to identify the *point estimate*, which comes from the sample data
5. Use the Standard Error Decision Tree to identify how to compute the *standard error* ($S_E$)
6. Identify or compute the appropriate *critical value*
7. Find the *lower-* and *upper-bounds* for the confidence interval by evaluating the "formula"
8. Interpret the confidence interval in the appropriate context

## Reminder: Hypothesis Testing Framework {.smaller}

**General Strategy:** Each time we conduct a hypothesis test, we will...

1. Read the scenario carefully and determine whether you are testing a claim about a mean, a proportion, comparison of means, or comparison of proportions.
2. Identify the claim being tested and state the *null hypothesis* ($H_0$) and *alternative hypothesis* ($H_a$). The null hypothesis represents the "status quo" or "no difference" and always involves an equal sign, while the alternative hypothesis is the claim to be tested and involves one of: $<,~>,~\neq$
3. Draw a picture of your alternative hypothesis, shading in the tails corresponding to samples that would satisfy the alternative hypothesis.
4. Set the level of significance ($\alpha$) for the test. The level of significance is the "cut-off" for a sample being unusual/unlikely/unexpected. The standard cutoff is $\alpha = 0.05$, unless we are told otherwise.
5. Compute the *test statistic*, where $\displaystyle{\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}}$

    + The *point estimate* comes from the sample data
    + The *null value* comes from the null hypothesis
    + The *standard error* ($S_E$) formula is found on the Standard Error Decision Tree

6. The test statistic is just a boundary value, use it along with your picture from step 3 to calculate a $p$-value (the probability of observing data at least as extreme as ours if the null hypothesis were true).
7. Compare your $p$-value to the level of significance ($\alpha$) demanded by your test and make a decision about whether or not to reject the null hypothesis.
8. Interpret the result of your hypothesis test in the context appropriate for your scenario.

## Inference on a Single Proportion versus Comparisons of Proportions

. . . 

At our last class meeting, we saw one example of the construction of a confidence interval and another example conducting a hypothesis test

. . . 

In both of those scenarios, the population parameter of interest was a single *population proportion*

. . . 

We'll see scenarios like this again today, but we'll also consider scenarios in which we'd like to compare two population proportions.

. . . 

The general procedures for constructing confidence intervals or conducting hypothesis tests are the same in both scenarios, except that in the comparison scenarios...

+ the population parameter of interest is the *difference in proportions* between the two sub-populations
+ the formula for computing the standard error ($S_E$) is different

## Navigating the Standard Error Tree

. . . 

One of the main purposes of the *Standard Error Decision Tree* document is to help you identify the appropriate standard error ($S_E$) to use in your calculation of a confidence interval or test statistic

. . . 

To navigate the tree, begin from the box at the top, which asks whether we are working with means ($\mu$) or proportions ($p$)

. . . 

Follow your answer down the tree and continue answering the questions until you land inside a box with a standard error formula ($S_E = \cdots$)

. . . 

The side of the tree for working with proportions is less complicated than the side for means -- we'll stick to proportions for now

. . . 

At this point, we'll either calculate $\displaystyle{S_E = \sqrt{\frac{p\left(1 - p\right)}{n}}}$ for a single proportion or $\displaystyle{S_E = \sqrt{\frac{p_1\left(1 - p_1\right)}{n_1} + \frac{p_2\left(1 - p_2\right)}{n_2}}}$ for a comparison between two population proportions

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

::::{.columns}

:::{.column width="50%"}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$

:::

:::{.column width="50%"}

:::

::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::{.columns}

::::{.column width="50%"}

:::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

:::

3. See the picture of the alternative hypothesis below:

::::

::::{.column width="50%"}

::::

:::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

:::::

4. Set $\alpha = 0.10$

::::::

::::::{.column width="50%"}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$


::::::

:::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

Notice that $\displaystyle{p_{\text{swipe}} = \frac{46}{200}}$, so $p_{\text{swipe}} = 0.23$ and<br/> $\displaystyle{p_{\text{personality}} = \frac{65}{180}}$, so $p_{\text{personality}} \approx 0.3611$

:::

:::::

::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

Notice that $\displaystyle{p_{\text{swipe}} = \frac{46}{200}}$, so $p_{\text{swipe}} = 0.23$ and<br/> $\displaystyle{p_{\text{personality}} = \frac{65}{180}}$, so $p_{\text{personality}} \approx 0.3611$

Notice also that our *standard error* is calculated as

$$S_E = \sqrt{\frac{p_{\text{swipe}}\left(1 - p_{\text{swipe}}\right)}{n_{\text{swipe}}} + \frac{p_{\text{personality}}\left(1 - p_{\text{personality}}\right)}{n_{\text{personality}}}}$$

:::

:::::

::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

Notice that $\displaystyle{p_{\text{swipe}} = \frac{46}{200}}$, so $p_{\text{swipe}} = 0.23$ and<br/> $\displaystyle{p_{\text{personality}} = \frac{65}{180}}$, so $p_{\text{personality}} \approx 0.3611$

Notice also that our *standard error* is calculated as

$$S_E = \sqrt{\frac{p_{\text{swipe}}\left(1 - p_{\text{swipe}}\right)}{n_{\text{swipe}}} + \frac{p_{\text{personality}}\left(1 - p_{\text{personality}}\right)}{n_{\text{personality}}}}$$

So, 

$$\texttt{test statistic} = \frac{\left(0.23 - 0.3611\right) - 0}{\sqrt{\frac{0.23\left(1 - 0.23\right)}{200} + \frac{0.3611\left(1 - 0.3611\right)}{180}}} \approx -2.82$$

:::

:::::

::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

$$\texttt{test statistic} \approx -2.82$$

:::

6. Now we'll draw our test statistic on our standard normal distribution and calculate the $p$-value

:::::

::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

$$\texttt{test statistic} \approx -2.82$$

6. Now we'll draw our test statistic on our standard normal distribution and calculate the $p$-value

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
ggplot() + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd = 1.5) + 
  geom_vline(xintercept = -2.82,
             lwd = 1.5,
             linetype = "dashed") + 
  scale_x_continuous(breaks = c(-2.82, -2, 0, 2)) + 
  labs(
    title = "",
    x = "",
    y = ""
  )
```

:::

:::{.column width="20%"}

:::

::::

:::::

7. $p$-value = ...

::::::

:::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

$$\texttt{test statistic} \approx -2.82$$

6. Now we'll draw our test statistic on our standard normal distribution and calculate the $p$-value

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
ggplot() + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd = 1.5) + 
  geom_vline(xintercept = -2.82,
             lwd = 1.5,
             linetype = "dashed") + 
  scale_x_continuous(breaks = c(-2.82, -2, 0, 2)) + 
  labs(
    title = "",
    x = "",
    y = ""
  )
```

:::

:::{.column width="20%"}

:::

::::

7. $p$-value = `2*pnorm(-2.82, 0, 1)` $\approx$ `r round(2*pnorm(-2.82, 0, 1), 4)`

:::::

::::::

:::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

$$\texttt{test statistic} \approx -2.82$$

6. Now we'll draw our test statistic on our standard normal distribution and calculate the $p$-value

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
ggplot() + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd = 1.5) + 
  geom_vline(xintercept = -2.82,
             lwd = 1.5,
             linetype = "dashed") + 
  scale_x_continuous(breaks = c(-2.82, -2, 0, 2)) + 
  labs(
    title = "",
    x = "",
    y = ""
  )
```

:::

:::{.column width="20%"}

:::

::::

7. $p$-value = `2*pnorm(-2.82, 0, 1)` $\approx$ `r round(2*pnorm(-2.82, 0, 1), 4)`

8. $p$-value $< \alpha$, we reject $H_0$ and accept $H_a$

:::::

::::::

:::::::

## A Completed Hypothesis Test {.smaller}

**Scenario:** A group of students is curious about whether the success rates (finding a long-term relationship) differ between people using a swipe-based dating app versus a personality-based matchmaking app. They conduct a survey of users from both types of apps and compare the proportion of users who report being in a long-term relationship. On the swipe-based app, 46 out of 200 users surveyed reported being in a long-term relationship. On the personality-based site, 65 out of 180 users surveyed said the same. Is there significant evidence to suggest (at the 10% level of significance) that proportion of successful relationships are not the same for swipe-based apps and personality-based apps?

:::::::{.columns}

::::::{.column width="50%"}

:::::{.nonincremental}

1. We're testing a claim comparing two population<br/> proportions
2. The hypotheses are: $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} = p_{\text{personality}}\\ H_a & : & p_{\text{swipe}} \neq p_{\text{personality}}\end{array}$, but a<br/> better way to write these them is $\begin{array}{lcl} H_0 & : & p_{\text{swipe}} - p_{\text{personality}} = 0\\ H_a & : & p_{\text{swipe}} - p_{\text{personality}} \neq 0\end{array}$

3. See the picture of the alternative hypothesis below:

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
x_vals <- seq(-3, 3, length.out = 500)
x_vals_shade <- seq(1.96, 3, length.out = 250)

ggplot() + 
  geom_ribbon(aes(x = x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_ribbon(aes(x = -x_vals_shade,
                  ymin = 0,
                  ymax = dnorm(-x_vals_shade, 0, 1)),
              fill = "steelblue") + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd =  1.5) + 
  scale_x_continuous(breaks = NULL) + 
  labs(title = "Samples Favoring Ha",
       x = "",
       y = "")
```

:::

:::{.column width="20%"}

:::

::::

4. Set $\alpha = 0.10$

:::::

::::::

::::::{.column width="50%"}

:::::{.nonincremental}

5. We'll calculate the test statistic, where $\texttt{test statistic} = \frac{\left(\texttt{point estimate}\right) - \left(\texttt{null value}\right)}{S_E}$

$$\texttt{test statistic} \approx -2.82$$

6. Now we'll draw our test statistic on our standard normal distribution and calculate the $p$-value

::::{.columns}

:::{.column width="20%"}

:::

:::{.column width="60%"}

```{r}
ggplot() + 
  geom_line(aes(x = x_vals,
                y = dnorm(x_vals, 0, 1)),
            lwd = 1.5) + 
  geom_vline(xintercept = -2.82,
             lwd = 1.5,
             linetype = "dashed") + 
  scale_x_continuous(breaks = c(-2.82, -2, 0, 2)) + 
  labs(
    title = "",
    x = "",
    y = ""
  )
```

:::

:::{.column width="20%"}

:::

::::

7. $p$-value = `2*pnorm(-2.82, 0, 1)` $\approx$ `r round(2*pnorm(-2.82, 0, 1), 4)`

8. Our observed data is incompatible with the proportion of successful relationships being equal. The relationship success rates differ across the two apps.

:::::

::::::

:::::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

:::

:::{.column width="50%"}

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

:::

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

:::

:::{.column width="50%"}

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

:::

:::

:::{.column width="50%"}

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

The *point estimate* is 

$$p_{\text{virtual}} - p_{\text{in-person}} \approx 0.1262$$

:::

:::

:::{.column width="50%"}

5. The *standard error* is...

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

The *point estimate* is 

$$p_{\text{virtual}} - p_{\text{in-person}} \approx 0.1262$$

:::

:::

:::{.column width="50%"}

:::{.nonincremental}

5. The *standard error* is...

\begin{align*}S_E &= \sqrt{\frac{p_{\text{virtual}}\left(1 - p_{\text{virtual}}\right)}{n_{\text{virtual}}} + \frac{p_{\text{in-person}}\left(1 - p_{\text{in-person}}\right)}{n_{\text{in-person}}}}\end{align*}

:::

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

The *point estimate* is 

$$p_{\text{virtual}} - p_{\text{in-person}} \approx 0.1262$$

:::

:::

:::{.column width="50%"}

:::{.nonincremental}

5. The *standard error* is...

\begin{align*} S_E &= \sqrt{\frac{p_{\text{virtual}}\left(1 - p_{\text{virtual}}\right)}{n_{\text{virtual}}} + \frac{p_{\text{in-person}}\left(1 - p_{\text{in-person}}\right)}{n_{\text{in-person}}}}\\
&\approx \sqrt{\frac{0.4429\left(1 - 0.4429\right)}{140} + \frac{0.3167\left(1 - 0.3167\right)}{120}}\end{align*}

:::

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

The *point estimate* is 

$$p_{\text{virtual}} - p_{\text{in-person}} \approx 0.1262$$

:::

:::

:::{.column width="50%"}

:::{.nonincremental}

5. The *standard error* is...

\begin{align*} S_E &= \sqrt{\frac{p_{\text{virtual}}\left(1 - p_{\text{virtual}}\right)}{n_{\text{virtual}}} + \frac{p_{\text{in-person}}\left(1 - p_{\text{in-person}}\right)}{n_{\text{in-person}}}}\\
&\approx \sqrt{\frac{0.4429\left(1 - 0.4429\right)}{140} + \frac{0.3167\left(1 - 0.3167\right)}{120}}\\
&\approx \sqrt{0.00176 + 0.0018}\end{align*}

:::

:::

::::

## A Completed Confidence Interval {.smaller}

**Scenario:** A college wants to understand if there is a difference in the proportion of students who attend office hours in-person versus virtually. Faculty members noticed varying attendance patterns and want to explore whether virtual office hours increase engagement. Out of 120 students surveyed, 38 said that they attend in-person office hours regularly. A separate random sample of 140 students revealed that 62 attended virtual office hours regularly. Construct a 90% confidence interval for the difference in proportions of students attending virtual and face-to-face office hours regularly.

::::{.columns}

:::{.column width="50%"}

:::{.nonincremental}

1. Read the problem carefully $\checkmark$
2. Estimate the *difference between the proportions* of students regularly attending virtual office hours versus in-person office hours
3. The "formula" for the confidence<br/> interval is 

$$\scriptsize{\left(\begin{array}{c}\texttt{point}\\ \texttt{estimate}\end{array}\right) \pm  \left(\begin{array}{c}\texttt{critical}\\ \texttt{value}\end{array}\right)\cdot \left(\begin{array}{c}\texttt{standard}\\ \texttt{error}\end{array}\right)}$$

4. The *point estimate* is the difference in proportions...

Notice that $\displaystyle{p_{\text{virtual}} = \frac{62}{140}}$ and $\displaystyle{p_{\text{in-person}} = \frac{38}{120}}$, so $p_{\text{virtual}} \approx 0.4429$ and $p_{\text{in-person}} \approx 0.3167$

The *point estimate* is 

$$p_{\text{virtual}} - p_{\text{in-person}} \approx 0.1262$$

:::

:::

:::{.column width="50%"}

:::{.nonincremental}

5. The *standard error* is...

\begin{align*} S_E &= \sqrt{\frac{p_{\text{virtual}}\left(1 - p_{\text{virtual}}\right)}{n_{\text{virtual}}} + \frac{p_{\text{in-person}}\left(1 - p_{\text{in-person}}\right)}{n_{\text{in-person}}}}\\
&\approx \sqrt{\frac{0.4429\left(1 - 0.4429\right)}{140} + \frac{0.3167\left(1 - 0.3167\right)}{120}}\\
&\approx \sqrt{0.00176 + 0.0018}\\
&\approx 0.0597\end{align*}

:::

6. The *critical value* is $\displaystyle{z_{\alpha/2} = 1.65}$

7. The confidence bounds are $\displaystyle{0.1262 \pm 1.65\left(0.0597\right) \to \left\{\begin{array}{lcl} 0.1262 + 0.0985 & = & 0.2247\\ 0.1262 - 0.0985 & = & 0.0277\end{array}\right.}$

8. We are 90% confident that the true difference in proportions of students regularly attending virtual office hours versus in person office hours is somewhere between 2.77% and 22.47%, with virtual office hour attendance being more prevalent.

:::

::::


## Examples to Try: Influencer and Viral Videos

**Scenario:** A social media influencer claims that 60% of their followers have watched their latest viral video. A marketing company surveys 400 randomly selected followers and finds that 54% have actually watched the video. Is there enough evidence to reject the influencer's claim at the 5% significance level?

## Examples to Try: Fast Food Preferences

**Scenario:** A fast food chain claims that 70% of customers prefer its chicken sandwich over its competitors'. A random sample of 300 customers shows that only 65% prefer the chain's sandwich. Construct a 95% confidence interval for the true proportion of customers who prefer the sandwich. Does the chain's claim fall within the confidence interval?

## Examples to Try: Vaccine Approval Rate

**Scenario:** In clinical trials, a new vaccine was found to be effective in 85% of cases. After the vaccine was approved, a public health survey randomly sampled 1,200 people who received the vaccine and found that 83% of them reported its effectiveness. Is this evidence that the true effectiveness rate of the vaccine is different from 85%? Test at the 1% significance level.

## Examples to Try: Coffee Shop Loyalty Cards

**Scenario:** A coffee shop chain claims that 40% of its customers use loyalty cards. A random sample of 250 customers reveals that 38% of them use the loyalty cards. At the 10% significance level, can you reject the coffee shop's claim?

## Examples to Try: Recycling Habits in a City

**Scenario:** A city's environmental department claims that 65% of its residents participate in curbside recycling programs. A local survey of 800 randomly selected residents finds that 68% of them recycle. Construct a 98% confidence interval for the true proportion of recyclers in the city. Does this support the department's claim?

## Examples to Try: Return Rates for an Online Store

**Scenario:** An online store states that its customer return rate is 8%. A customer service audit randomly selects 700 transactions and finds that 6.5% of the customers returned items. Construct a 95% confidence interval for the true return rate. Is the store’s stated return rate within this interval?

## Examples to Try: Video Game Console Popularity

**Scenario:** A gaming company believes that 75% of video game players prefer their console over competitors'. A survey of 500 gamers reveals that 72% prefer the company’s console. Construct a 98% confidence interval for the true proportion of gamers who prefer the console. Does the company's claim fall within this interval?

## Examples to Try: Support for a New Environmental Policy

**Scenario:** A politician claims that a minority of the public supports a new environmental policy. A random sample of 900 people finds that 54% of them support the policy. At the 5% significance level, can you reject the politician's claim?

## Next Time...

<center><br/> Practicing with Inference</center>



