---
title: "Reading Output from Statistical Software"
author: Dr. Gilbert
format: revealjs
date: today
date-format: long
theme: serif
incremental: true
fontsize: 20pt
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(patchwork)
library(countdown)
library(statsr)

options(kable_styling_bootstrap_options = c("hover", "striped"))
options(scipen = 999)

theme_set(theme_bw(base_size = 32))

set.seed(123)

pop_prop <- 0.86
num_obs <- 17
new_obs <- 113
num_samps <- 100
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}

div.cell-output-stdout {
  font-size: 1.5em;
}

a {
  color: purple;
}

a:link {
  color: purple;
}

a:visited {
  color: purple;
}
```

## The Highlights

. . .

It is often the case that practitioners will utilize software to conduct statistical inference

. . . 

We'll just focus on navigating and interpreting the software output in these slides

. . .

After completing several examples, we'll return to our long list of practice scenarios

## Scenario I: Average Highway Fuel Economy

. . .

**Scenario:** Researchers are interested in understanding the fuel efficiency of cars on the highway. To investigate, they collect data on the highway gas mileage (in miles per gallon) of a random sample of 234 cars. The researchers want to determine whether the average highway gas mileage for all cars exceeds 23 mpg. They conduct a test at the $\alpha = 0.10$ level of significance and the results appear below. Write the hypotheses for the test and determine the results of the test in the context of the scenario.

. . . 

```{r}
inference(y = hwy, data = mpg, type = "ht", statistic = "mean", method = "theoretical", null = 23, alternative = "greater", show_eda_plot = FALSE, show_inf_plot = FALSE)
```
## Scenario II: Comparing Average Highway Fuel Economy

**Scenario:** The researchers from the previous scenario would like to determine whether the highway fuel economy of automatic and manual vehicles differ from on another. They use their random sample of 234 cars and conduct this test at the $\alpha = 0.05$ level of significance. The results of the test appear below. Write the hypotheses for the test and state the conclusion in the context of the scenario.

. . .

```{r}
mpg <- mpg %>%
  mutate(trans = str_sub(trans, 1, -5))

inference(y = hwy, x = trans, data = mpg, type = "ht", statistic = "mean", method = "theoretical", null = 0, alternative = "twosided", show_eda_plot = FALSE, show_inf_plot = FALSE)
```

## Scenario III: City Gas Mileage for Four-Wheel Drive Vehicles

**Scenario:** The researchers are now interested in estimating the city gas mileage for four-wheel drive vehicles. Using the four-wheel drive vehicles from their random sample, then construct a 95% confidence interval for city gas mileage. The results appear below. Interpret them in the context of the scenario.

. . . 

```{r}
four_wd <- mpg %>%
  filter(drv == "4")

inference(y = cty, data = four_wd, type = "ci", statistic = "mean", method = "theoretical", conf_level = 0.95, show_eda_plot = FALSE, show_inf_plot = FALSE)
```