---
title: "Reading Output from Statistical Software"
author: Dr. Gilbert
format: revealjs
date: today
date-format: long
theme: serif
incremental: true
fontsize: 20pt
---

```{r global-options, include=FALSE}
library(tidyverse)
library(tidymodels)
library(patchwork)
library(countdown)
library(statsr)

options(kable_styling_bootstrap_options = c("hover", "striped"))
options(scipen = 999)

theme_set(theme_bw(base_size = 32))


```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}

div.cell-output-stdout {
  font-size: 1.5em;
}

a {
  color: purple;
}

a:link {
  color: purple;
}

a:visited {
  color: purple;
}
```

## The Highlights

. . .

It is often the case that practitioners will utilize software to conduct statistical inference

. . . 

We'll just focus on navigating and interpreting the software output in these slides

. . .

After completing several examples, we'll return to our long list of practice scenarios

## Scenario I: Average Highway Fuel Economy

. . .

**Scenario:** Researchers are interested in understanding the fuel efficiency of cars on the highway. To investigate, they collect data on the highway gas mileage (in miles per gallon) of a random sample of 234 cars. The researchers want to determine whether the average highway gas mileage for all cars exceeds 22.5 mpg. They conduct a test at the $\alpha = 0.10$ level of significance and the results appear below. Write the hypotheses for the test and determine the results of the test in the context of the scenario.

. . . 

```{r}
inference(y = hwy, data = mpg, type = "ht", statistic = "mean", method = "theoretical", null = 22.5, alternative = "greater", show_eda_plot = FALSE, show_inf_plot = FALSE)
```

. . . 

FYI...

. . .

...the code to run this inference is

. . .

```{r}
#| echo: true
#| eval: false

inference(y = hwy, data = mpg, type = "ht", 
          statistic = "mean", method = "theoretical", 
          null = 22.5, alternative = "greater", 
          show_eda_plot = FALSE, show_inf_plot = FALSE)
```

## Scenario II: Comparing Transmission Types

**Scenario:** The researchers from the previous scenario would like to determine whether the proportion of automatic and manual vehicles differ from one another between front-wheel and rear-wheel drive vehicles. They use their random sample of 234 cars and conduct this test at the $\alpha = 0.05$ level of significance. The results of the test appear below. Write the hypotheses for the test and state the conclusion in the context of the scenario.

. . .

```{r}
drive_trans <- mpg %>%
  mutate(trans = str_sub(trans, 1, -5)) %>%
  filter(drv %in% c("f", "r")) %>%
  mutate(drv = ifelse(drv == "f", "front", "rear"))

inference(y = trans, x = drv, data = drive_trans, 
          type = "ht", statistic = "proportion", 
          method = "theoretical", null = 0,
          alternative = "twosided", 
          success = "manual",
          show_eda_plot = FALSE, 
          show_inf_plot = FALSE)
```

. . . 

FYI...

. . .

...after a bit of data manipulation, the code to run this inference is

. . .

```{r}
#| echo: true
#| eval: false

inference(y = trans, x = drv, data = drive_trans, 
          type = "ht", statistic = "proportion", 
          method = "theoretical", null = 0,
          alternative = "twosided", 
          success = "manual",
          show_eda_plot = FALSE, 
          show_inf_plot = FALSE)
```

## Scenario III: City Gas Mileage for Four-Wheel Drive Vehicles

**Scenario:** The researchers are now interested in estimating the city gas mileage for four-wheel drive vehicles. Using the four-wheel drive vehicles from their random sample, then construct a 95% confidence interval for city gas mileage. The results appear below. Interpret them in the context of the scenario.

. . . 

```{r}
four_wd <- mpg %>%
  filter(drv == "4")

inference(y = cty, data = four_wd, type = "ci", statistic = "mean", method = "theoretical", conf_level = 0.95, show_eda_plot = FALSE, show_inf_plot = FALSE)
```

. . . 

FYI...

. . .

...after a bit of data manipulation, the code to run this inference is

. . .

```{r}
#| echo: true
#| eval: false

inference(y = cty, data = four_wd, type = "ci", 
          statistic = "mean", method = "theoretical", 
          conf_level = 0.95, show_eda_plot = FALSE, 
          show_inf_plot = FALSE)
```

## Scenario IV: Difference in Highway and City Fuel Economy

**Scenario:** As one final investigation, the researchers would like to estimate the difference in highway gas mileage and city gas mileage. They build a 90% confidence interval for the difference in gas mileage and the results appear below. Interpret the results in context.

. . . 

```{r}
cty_hwy_mpg <- mpg %>%
  select(cty, hwy) %>%
  pivot_longer(cols = everything(), names_to = "environment", values_to = "mpg") %>%
  mutate(environment = factor(environment, levels = c("hwy", "cty")))

inference(y = mpg, x = environment, 
          data = cty_hwy_mpg, type = "ci", 
          statistic = "mean", method = "theoretical", 
          conf_level = 0.90, show_eda_plot = FALSE, 
          show_inf_plot = FALSE)
```

. . . 

FYI...

. . .

...after a bit of data manipulation, the code to run this inference is

. . .

```{r}
#| echo: true
#| eval: false

inference(y = mpg, x = environment, 
          data = cty_hwy_mpg, type = "ci", 
          statistic = "mean", method = "theoretical", 
          conf_level = 0.90, show_eda_plot = FALSE, 
          show_inf_plot = FALSE)
```

