---
title: "Youth Behavior and Risk"
author: "Me, Analyst"
date: today
date-format: medium
format: html
theme: cerulean
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

### Overview

In this notebook, we'll develop and test claims on multiple proportions. In particular, we'll test a *Goodness of Fit* claim and also a claim about *dependence/independence*. I had planned to use the Youth Risk Behavioral Surveillance System data for this, but we can use the Minnesota Police Use of Force data instead if we prefer.

## Our Data Set

We'll decide on our data set and read it in below.

```{r}
#Read in our chosen data set using this code cell...

```

We'll describe our data set briefly below.

### Our Research Questions and Hypotheses

We'll define our research questions and hypotheses below.

### Statistical Inference

We'll conduct our inference here.

```{r}
#We'll use this code cell (and perhaps some others) for conducting our analysis and making inference...
```

***

## Summary

We'll summarize the work we've done here.