---
title: "Youth Behavior and Risk"
author: "Me, Analyst"
date: today
date-format: medium
format: html
theme: cerulean
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

### Overview

In this notebook, we'll develop and test claims on multiple population means. In particular, we'll test *Analysis of Variance* (ANOVA) to test whether body measurements differ across three different species of penguin.

## Our Data Set

```{r message = FALSE, warning = FALSE}
penguins <- read_csv("https://raw.githubusercontent.com/mwaskom/seaborn-data/master/penguins.csv")
```

We'll describe our data set briefly below.

### Our Research Questions and Hypotheses

We'll define our research questions and hypotheses below.

### Statistical Inference

We'll conduct our inference here.

```{r}
#We'll use this code cell (and perhaps some others) for conducting our analysis and making inference...
```

***

## Summary

We'll summarize the work we've done here.