---
title: "Activity: Inference with Raw Data"
author: "MAT241 Class"
format: 
  html:
    theme: flatly
  pdf: default
toc: true
---

```{r setup, include = FALSE}
library(tidyverse)
library(kableExtra)
library(nflfastR)
library(taylor)
library(palmerpenguins)
```

:::{.callout-note}

## Notebook Purpose

In this notebook, we'll have an option to investigate inferential questions with raw data. We'll propose our own questions and choose to focus on data that we find interesting.

:::

:::{.callout-important}

## Available Data Sets

There are lots of freely accessible data sets out there. You just need to know where to find them. For this activity, I've selected just a few so that we don't get bogged down in identifying data.

+ Taylor Swift Songs
+ NFL Play By Play Data
+ Penguin Anatomy and Physiology Data

:::

### Taylor Swift

```{r}
taylor_album_songs %>%
  select(-lyrics) %>%
  head() %>%
  kable()
```

### NFL Play By Play

```{r}
pbp_2324 <- load_pbp(2023:2024)

set.seed(2324)
pbp_2324_samp <- pbp_2324 %>%
  filter(play_type %in% c("pass", "run")) %>%
  sample_n(size = 5e3)

pbp_2324_samp %>%
  select(-desc, -weather, -nfl_api_id) %>%
  head() %>%
  kable()
```


### Penguin Anatomy and Physiology

```{r}
penguins %>%
  head() %>%
  kable()
```




