---
title: "Intro to R with Austin Housing Data"
author: Dr. Gilbert
format: 
  revealjs:
    fontsize: 20pt
date: today
date-format: long
theme: serif
incremental: true
---

```{r global-options, include=FALSE}
library(tidyverse)
library(reticulate)

theme_set(theme_bw(base_size = 20))
```

```{python}
import numpy as np
import matplotlib.pyplot as plt
```

```{css}
code.sourceCode {
  font-size: 1.3em;
  /* or try font-size: xx-large; */
}

a {
  color: purple;
}

a:link {
  color: purple;
}

a:visited {
  color: purple;
}
```

```{r}
library(tidyverse)
library(countdown)

austin <- read_csv("https://raw.githubusercontent.com/agmath/agmath.github.io/master/data/classification/austin_zillow_data.csv")
```

## Starter File

1. Create an empty Quarto Document in RStudio by navigating to `File -> New File -> Quarto Document...` using the menus.
2. Choose to *Create Empty Document` and then remove the minimal YAML header to start with a truly empty document.
3. Navigate to [the starter notebook here](https://github.com/agmath/IntroductoryStatistics/blob/main/Days2to5_AustinHousingData.qmd).
4. Copy all of the lines in that document and paste them into your Quarto document, making sure that you are in the *Source* editor rather than the *Visual* editor.
5. Add to that file as we work through these slides together.

## Review of Quarto

. . . 

We work in Quarto documents, which allow a mixture of code and text in a single document

. . . 

Quarto documents consist largely of the following components

1. The YAML header governs global document properties and appears at the top of the document between *code fences* (`---`)
2. Code chunks begin and end with triple backticks (see below), creating sections of the document over a *grey background*
3. Sections of the document over a white background are just text cells and you can type freely there -- if you want to try some markdown to format your text, then you are free to do so!

. . .

An example of a code chunk

```{{r}}
#Code goes here...

```

## YAML Overview

. . .

We'll use very simple YAML headers in this course, but you are welcome to explore more complex document customization if you like. 

. . .

Our YAML header will generally look like the following:

```{r}
#| echo: true
#| eval: false

---
title: "Notebook Title"
author: "Your Name"
format: html
date: today
date-format: long
theme: flatly
---
```

. . . 

You can actually just use the `title`, `author`, and `format` settings if you prefer.

+ **Options for `format`:** We'll use `html` by default, but you can use `docx` to output a Word Document or `pdf` to output a PDF file
+ **Options for `theme`:** You can [find the available document themes here](https://quarto.org/docs/output-formats/html-themes.html)

## Basic Markdown Overview

. . . 

Markdown formatting in this course is optional, but using markdown can make your documents look quite nice. The following are the most common pieces of markdown you might find use for.

+ Headings are built using the *hashtag* symbol (\#) -- more hashtags means a lower-level (smaller font) heading
+ We can surround text by a single asterisk to make that text italics -- for example \*this is italics\* formats as *this is italics*
+ We can surround text by double asterisks to make that text bold -- for example \*\*this is bold\*\* formats as **this is bold**
+ We can surround text by backticks to format that text as "code" -- for example \`this is code\` formats as `this is code`
+ Bulleted lists, like this one can be made by starting each line with a "+" sign (you'll need an empty line preceeding the list)
+ Numbered lists are the same as bulleted lists, but you just start each line with a number

## Review of R So Far

. . . 

Within a Quarto Document, R code is run inside of a code chunk like the one below:

```{{r}}

````

+ By default, R can do basic calculations

  + We load libraries to enable specialized functionality -- for example, by running `library(tidyverse)` in a code chunk
  
+ We run R code by holding `ctrl` and hitting `Enter` or `Return`

+ We store items in variables using the arrow operator -- for example,<br/> by running `x <- 2` in a code chunk

  + We can print out the contents of an object/variable just by calling its name -- for example running `x` in a code chunk would print `2`

+ We can read data from a *csv* file using `read_csv("file_path")`

  + Don't forget to store the result into a variable if you want to use the data later!


## Reminder of the Austin, TX Housing Data

A *data dictionary* appears below:

+ `id` is a row number (unique identifier)
+ `description` is a free-form text field, describing the property (unique identifier...for our course)
+ `city`, `homeType`, `hasSpa`, and `priceRange` are all categorical variables
+ `latitude`, `longitude`, `lotSizeSqFt`, `avgSchoolRating`, and `MedianStudentsPerTeacher` are all numerical variables
+ `garageSpaces`, `yearBuilt`, `numOfPatioAndPorchFeatures`, `numOfBathrooms`, and `numOfBedrooms` could be treated as either numerical or categorical variables

  + We get to choose how to treat them

## In-Class Activity, Part I

. . .

$\bigstar$ Work with the people next to you to come up with some questions which would be interesting to investigate with our Austin housing data.

+ Start with some "easy" questions about single variables at a time
+ Move to some questions about combinations of variables

. . . 

```{r}
countdown(minutes = 5)
```

## In-Class Activity, Part I

$\bigstar$ Work with the people next to you to come up with some questions which would be interesting to investigate with our Austin housing data.

:::{.nonincremental}

+ Start with some "easy" questions about single variables at a time
+ Move to some questions about combinations of variables

:::

$\bigstar$ What questions did we come up with?

```{r}
countdown(minutes = 7)
```

## In-Class Activity, Part I

$\bigstar$ Work with the people next to you to come up with some questions which would be interesting to investigate with our Austin housing data.

:::{.nonincremental}

+ Start with some "easy" questions about single variables at a time
+ Move to some questions about combinations of variables

:::

$\bigstar$ What questions did we come up with?

$\bigstar$ Take a few minutes to write those questions into your `Day2to5_AustinHousingData.qmd` file. Render your notebook to make sure everything looks the way you intended. Try some markdown formatting to improve the structure and readability of your notebook.

```{r}
countdown(minutes = 10)
```

## In-Class Activity, Part II

. . .

$\bigstar$ Work with the people next to you to decide which of your questions are just about your *sample data* and which of your questions are about the entire *population*.

```{r}
countdown(minutes = 2)
```

## In-Class Activity, Part II

$\bigstar$ Work with the people next to you to decide which of your questions are just about your *sample data* and which of your questions are about the entire *population*.

$\bigstar$ Update your notebook to include two subsections -- one with *sample*-level questions and the other with *population*-level questions. When finished, you should have two versions of every one of the questions you initially wrote down.

```{r}
countdown(minutes = 5)
```

## In-Class Activity, Part II

$\bigstar$ Work with the people next to you to decide which of your questions are just about your *sample data* and which of your questions are about the entire *population*.

$\bigstar$ Update your notebook to include two subsections -- one with *sample*-level questions and the other with *population*-level questions. When finished, you should have two versions of every one of the questions you initially wrote down.

$\bigstar$ What is the main difference in phrasing between *descriptive* (sample-level) questions and *inferential* (population-level) questions?

```{r}
countdown(minutes = 1)
```

## In-Class Activity, Part III

. . .

$\bigstar$ If we are going to use our available data to answer *inferential* (population-level) questions, then what assumption(s) are we making?

```{r}
countdown(minutes = 1)
```

## In-Class Activity, Part III

$\bigstar$ If we are going to use our available data to answer *inferential* (population-level) questions, then what assumption(s) are we making?

$\bigstar$ Can both types of question (*descriptive* and *inferential*) be answered simply by calculating summary statistics from our sample data? Why or why not?

```{r}
countdown(minutes = 1)
```

## In-Class Activity, Part III

$\bigstar$ If we are going to use our available data to answer *inferential* (population-level) questions, then what assumption(s) are we making?

$\bigstar$ Can both types of question (*descriptive* and *inferential*) be answered simply by calculating summary statistics from our sample data? Why or why not?

$\bigstar$ Without using R code just yet, describe what you would need to do in order to answer each of your *descriptive* questions. Add those descriptions to your notebook.

```{r}
countdown(minutes = 5)
```

## Closing

1. Render your notebook and make sure that the sections we've updated look as you intended them to.

    + Make any updates you like and re-render the notebook.

1. We didn't *really* use any R today, but we'll pick up where we left off next time and actually use R to answer the *descriptive* questions we've posed.

1. **Homework:** Complete and submit the `Topic 3` notebook at least 30 minutes before Monday's class meeting. That notebook will give you many of the tools we'll need for Monday.

. . . 

**Question:** Moving forward, should we continue using slide decks like this one?